Accelerated Magnetic Resonance Imaging by Adversarial Neural Network

### 2018-02-01 paper.tex
Tammy update - 01/02/2018

We compare the proposed method to reconstruction results obtained by using a conventional compressed sensing method CS-MRI \cite{lustig2007sparse} and Zero-filling. In addition, we optimize two networks: 1. CNN-L2 - a k-space generator (with same architecture as the proposed method) trained with only L2 loss. 2. IM-CNN-L2 - a network which performs on the image-space and optimized to remove the artifacts caused by under-sampling and zero padding.
To assess the clinical usability of the reconstructed images we perform an extended validation, composed of anatomy and quantitative measurements. In addition to the common used PSNR and Structure Similarity Index (SSIM) we perform tissue segmentation and compared the results to those obtained by using the original fully-sampled images.
Segmentation compatibility, measured in terms of Dice scores and Hausdorff distances, demonstrate the quality of the proposed MRI reconstruction with respect to other methods. The proposed method outperforms the other in 3T and 1.5T brain MRI reconstruction, using a large publicly available dataset, where only 40\%, 25\% and 16.6\% of the raw samples of each scan are used. Moreover, compared to the widely used CS-MRI, the proposed method is faster in a two orders of magnitude. This makes it possible to use it in a real time manner.

The paper is organized as follows. Section \ref{method_section} presents some theoretical foundation and our method. Section \ref{experiments_section} describes the experimental results. Conclusions and future directions are describes in section \ref{conclusions_section}.

@@

\subsection{K-space}
Let $\bm{u}$ denote the desired signal, a 2D MR image, obtained by the IFFT of the complex k-space signal $s_{0}$. Let $M_{F}$ denote a full sampling mask such that the reconstructed MR image is: 
\begin{equation}
\bm{u}=F^{H}M_{F}\odot s_{0}
\end{equation} 
where $H$ is the Hermitian transpose operation, $\odot$ denotes element-wise multiplication, and $F^{H}$ is an orthonormal 2D IFFT operator, such that $F^{H}F=I$. 
While sampling part of the k-space, using $M_{p}$ as a sampling mask, the reconstructed MR image suffers from artifacts and aliasing. An example artifacts caused by under-sampling and zero padding in the phase axis using Gaussian mask (40\%) is shown in Figure \textbf{\ref{fig:pair_of_k_space_mri}}.

@@

The training data consists of 500 $3D$ brain MRI (T1) scans of different patients from the IXI dataset\footnote{\url{http://brain-development.org/ixi-dataset/}}. The data has been acquired by three MR machines, Philips 1.5T, 3T and GE 1.5T. All images padded to resolution of $256\times256$ pixels.
From each $3D$ volume we extract 93 $2D$ saggital slices. We used $37.2k$ $(80\%)$ 2D slices for training and $9.2k$ $(20\%)$ for testing (100 $3D$ volumes).
In order to create k-space images for training, inverse orthonormal
2D FFT is applied to the fully-sampled MR images. 
Further to our work~\cite{shitrit2017accelerated}, which was been focused on 1D Cartesian sampling mask along the phase axis, in this work, we sample the k-space using $2D$ random Gaussian mask. By that, we exploit the sparsity in both phase and frequency dimensions and achieve higher sampling factors of 2.5, 4 and 6 (Figure~\ref{fig:sampling_masks}).
Data augmentation is created by random offsets of the proposed mask and image flipping. This leads us to reconstruction of the MR image from 40\%, 25\% and 16.6\% of the original k-space data.

The generator is composed of $5$ blocks of CONV-BatchNorm-ReLU, with output channels $16,32,64,32,8,$ respectively. The last layer is CONV with two outputs channels (for real and imaginary parts). The discriminator is composed of $4$ blocks of CONV-Pool-BatchNorm-LReLU with output channels $8,16,32,16$ and one fully-connected layer. All CONV layers kernel size is $3\times3$. All weights was initialized by Xavier \cite{glorot2010understanding}. We used RMSprop solver with fixed learning rate of $5\mathrm{e}{-6}$ and set $k_{d}$ to 1.

We compare the proposed method to reconstruction results obtained by using a conventional compressed sensing method CS-MRI \cite{lustig2007sparse} and Zero-filling. In addition, we optimized two networks: 1. CNN-L2 - a k-space generator (with same architecture as the proposed method) trained with only L2 loss. 2. IM-CNN-L2 - a network which performs on the image-space and optimized to remove the artifacts caused by under-sampling and zero padding. 
%We used the U-net \cite{ronneberger2015u} architecture inspired by \cite{lee2017deep}. 
We used the U-net \cite{ronneberger2015u} architecture and the same reconstruction method as presented in~\cite{hyun2017deep}. Specifically, we apply an FFT on the U-net output image to get an estimated k-space. The fully k-space is then recovered by integrating the sampled k-space according to the sampling mask. The same sampling masks was used for all methods.

>>>>

We compare the proposed method reconstructions to those obtained by using a conventional compressed sensing method CS-MRI \cite{lustig2007sparse} and Zero-filling. In addition, for comparison, we test two other networks: 1. CNN-L2 - a k-space generator (with same architecture as the proposed method) trained with only L2 loss. 2. IM-CNN-L2 - a network which performs on the image-space and is optimized to remove the artifacts caused by under-sampling and zero padding.
The method is applied to the Information eXtraction from Images (IXI) dataset, which is a large publicly available dataset, including 3T and 1.5T scans. Raw data is down-sampled such that only 40\%, 25\% and 16.6\% of the raw samples of each scan are used.
To assess the clinical usability of the reconstructed images we perform an extended validation, composed of anatomy and quantitative measurements. In addition to the common used PSNR and Structure Similarity Index (SSIM) we perform tissue segmentation and compare the results to those obtained by using the original fully-sampled images.
Segmentation compatibility, measured in terms of Dice scores and Hausdorff distances, demonstrate the quality of the proposed MRI reconstruction with respect to other methods.
The proposed method outperforms all other tested methods in all evaluation metrics. Moreover, compared to the widely used CS-MRI, the proposed method is faster in two orders of magnitude. In fact, it can work in real time, and is therefore practical for routine uses.

The rest of the paper is organized as follows. Section~\ref{method_section} provides a brief background on MRI acquisition and introduce the network architecture as well the training methodology. Section~\ref{experiments_section} presents the experimental results. Conclusions and future directions are given in section~\ref{conclusions_section}.

@@

\subsection{K-space}
Let $\bm{u}$ denote the desired signal, a 2D MR image, obtained by the IFFT of the complex k-space signal $s_{0}$. Let $M_{F}$ denote a full sampling mask such that the reconstructed MR image is: 
\begin{equation}
\bm{u}=F^{H}M_{F}\odot s_{0}
\end{equation} 
where $H$ is the Hermitian transpose operation, $\odot$ denotes element-wise multiplication, and $F^{H}$ is an orthonormal 2D IFFT operator, such that $F^{H}F=I$. 
While sampling part of the k-space, using $M_{p}$ as a sampling mask, the reconstructed MR image suffers from artifacts and aliasing. An example artifacts caused by under-sampling and zero padding in the phase axis using a Gaussian mask (40\%) is shown in Figure \textbf{\ref{fig:pair_of_k_space_mri}}.

@@

The training data consists of 500 $3D$ brain MRI (T1-weighed) scans of different patients from the IXI dataset\footnote{\url{http://brain-development.org/ixi-dataset/}}. The data has been acquired by three MR machines, Philips 1.5T, 3T and GE 1.5T using different imaging protocols.
From each $3D$ volume we extract 93 $2D$ saggital slices. All slices were interpolated to $256\times256$ pixels. We used $37.2k$ $(80\%)$ 2D slices for training and $9.3k$ $(20\%)$ for testing (100 $3D$ volumes).
In order to create k-space images for training, inverse orthonormal
2D FFT is applied to the fully-sampled MR images. 
Different from our preliminary work~\cite{shitrit2017accelerated}, which employed 1D Cartesian sampling mask along the phase axis, here, we perform k-space sampling using $2D$ random Gaussian mask. By this, we exploit the sparsity in both phase and frequency dimensions and therefore able to accommodate higher sampling factors (Figure~\ref{fig:sampling_masks}). MRI reconstruction is performed using 40\%, 25\% and 16.6\% of the original k-space data.
%Data augmentation is created by random offsets of raw images (prior to the IFFT) and k-space flipping. This leads us to reconstruction of the MR image from 40\%, 25\% and 16.6\% of the original k-space data.

The generator is composed of $5$ blocks of CONV-BatchNorm-ReLU, with output channels $16,32,64,32,8,$ respectively (see Figure~\ref{fig:architectures}). The last CONV layer is composed of two output channels for real and imaginary components. The discriminator is composed of $4$ blocks of CONV-Pool-BatchNorm-LReLU with output channels $8,16,32,16$ and one fully-connected layer. All CONV layer's kernel size is $3\times3$. All weights was initialized by Xavier \cite{glorot2010understanding}. We used RMSprop~\cite{tieleman2012lecture} solver with fixed learning rate of $5\mathrm{e}{-6}$ and set $k_{d}$ to 1.

We compared the proposed method reconstructions to those obtained by using a conventional compressed sensing method CS-MRI \cite{lustig2007sparse} and Zero-filling. In addition, we tested two other networks. To asses the use of adversarial loss, we trained a k-space generator, with same architecture as the proposed method, with only L2 loss (CNN-L2). We also implemented the method proposed in~\cite{hyun2017deep} and constructed a network IM-CNN-L2 which is applied directly to the image space rather than k-space. Specifically, we used the U-net architecture~\cite{ronneberger2015u} and applied an FFT on each of its output images to get an estimated k-space. The complete k-space is then recovered by integrating estimated samples with the sampled raw data according to the sampling mask. The same sampling masks were used for all methods.
